---
title: "week10sgta"
author: "Al"
date: "2023-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1
## Question 1
### a
```{r}
turtles <- read.csv("SGTADATASETS/turtles.csv", header = TRUE)
turt.1 <- lm(eggs ~ carapace, data = turtles)
summary(turt.1)

par(mfrow = c(1, 3))
plot(turtles)
abline(turt.1)
plot(turt.1, which = 1:2)
```

- The scatterplot shows a random distribution without much evidence of a linear relationship. However there does seem to be some quadratic pattern. 
- The residuals vs fitted plot shows definite curvature (concave down), suggesting lack of linearity assumption.
- There is skew in the qqplot suggesting residuals not very close to normally distributed.

### b
```{r}
turt.2 <- lm(eggs ~ carapace + I(carapace^2), data = turtles)
turt.3 <- lm(eggs ~ carapace + I(carapace^2) + I(carapace^3), data = turtles)
```

### c
```{r}
anova(turt.2)
anova(turt.3)
```
- The cubic term is insignificant, however the quadratic term is significant.

### d
```{r}
par(mfrow = c(1,2))
plot(turt.2, which = 1:2)

plot(turt.1, which = 1)
```
- Much better qqplot suggesting normality assumption.
- Spike in residuals vs fitted plot however points are mostly random suggesting quadratic fit is sufficient.

### e
```{r}
plot(turtles)
x <- seq(from = min(turtles$carapace), to = max(turtles$carapace), length = 32)
carapacedat <- data.frame(carapace = x)
eggshat <- predict(turt.2, newdata = carapacedat)
lines(x, eggshat)
```

## Question 2
### a
```{r}
sexdata <- read.table("SGTADATASETS/sexdiscrim.txt", header = TRUE)
sexdata$SEX <- factor(sexdata$SEX)
mycols <- c("blue", "red")
pairs(sexdata[, -2], panel = panel.smooth,
col = mycols[unclass(sexdata$SEX)])
```
In general men have higher BSALs than women. However hard to analyse plots as there is a lot of noise.

### b
```{r}
sex.1 <- lm(BSAL ~ AGE + EDUC + EXPER + SEX, data = sexdata)
summary(sex.1)

plot(sex.1, which = 1:2)
plot(resid(sex.1) ~ AGE + EDUC + EXPER, data = sexdata)
```
- Quantile plot of residuals appears to have a concave up shape, indicating skewness. 
- Variance seems fairly constant over fitted values.

### c
```{r}
sexdata$lnbsal <- log(sexdata$BSAL)
sex.2 <- lm(lnbsal ~ AGE + EDUC + EXPER + SEX, data = sexdata)
summary(sex.2)

plot(sex.2, which = 1:2)
plot(resid(sex.2) ~ AGE + EDUC + EXPER, data = sexdata)
```

### d
The quantile plot of residuals looks more linear, suggesting a better adherence to the model require-
ments. There is still curvature in the residual vs EXPER and AGE predictors. Could possibly add
quadratic term. In both the original model and log transformed model, EXPER is insignificant and
should be dropped.

### e
Both models are similar in terms of performance (R2) They both show significant predictors in the sex
and education levels. The experience levels are not significant. The log transformed model is superior
since it adheres to the linear regression framework and thus inferences used will be more reliable.

### f
```{r}
sex.4 <- update(sex.2, . ~ . - EXPER)
summary(sex.4)
```
```{r}
sex.3 <- update(sex.1, . ~ . - EXPER)
summary(sex.3)
```
### g
We are 95% confident that the starting salary is between 506.506 and 1024.4812 units higher
for males than females, on average, for the normal model. 

We are 95% confident that the starting salary is between 9.7% and 20.5% higher for males
than females, on average, for the log transformed model.




