---
title: "week9sgta"
author: "Al"
date: "2023-10-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1
## Question 1
### a
```{r}
aisfemales <- read.csv("SGTADATASETS/aisfemales.csv", header = TRUE)
plot(aisfemales)
#pairs(aisfemales, panel = panel.smooth)
```
LBM is highly correlated with Wt, less so but still highly correlated with Ht, with a high correlation
between Ht and Wt. Only slight correlation between LBM and RCC and Hg. High correlation
between Hg and RCC but not with other predictors

### b
```{r}
ais.1 <- lm(LBM ~ Ht + Wt + RCC + Hg, data = aisfemales)
summary(ais.1)
```
```{r}
par(mfrow = c(1, 2))
plot(ais.1, which = 1:2)
```
```{r}
par(mfrow = c(2, 2))
plot(resid(ais.1) ~ Ht + Wt + RCC + Hg, data = aisfemales)
```
1. The Normal Q-Q plot of residuals has a slight curvature but close to linear implying errors close to
normally distributed.
2. The residuals vs fitted has no discernible pattern.
3. Residuals vs predictor plots no obvious pattern. So the linear model seems adequate

### c
```{r}
ais.2 <- lm(LBM ~ Ht + Wt + Hg, data = aisfemales)
summary(ais.2)
```
From output we can see Hg has increased in significance from a P-Value from 0.0364 to 0.00487.
There was correlated information in RCC that has been removed that Hg is now providing (hence more
significant)

```{r}
par(mfrow = c(1, 2))
plot(ais.2, which = 1:2)
```
```{r}
par(mfrow = c(1, 3))
plot(resid(ais.2) ~ Ht + Wt + Hg, data = aisfemales)
```
Quantile plot looks linear and residual plots have no pattern, suggesting linear model still adequate.
The loss in R2 is only a percent, suggesting reduced model is better.

### d
```{r}
anova(ais.1)
anova(ais.2)
```
SS(ais.1) - SS(ais.2) = 1.27.

```{r}
anova(aov(LBM ~ Ht + Wt + Hg + RCC, data = aisfemales))
```


### e
```{r}
summary(ais.2)$coefficients
```
(0.4523, 0.5873). 

## Question 2
```{r}
lifeexp <- read.table("SGTADATASETS/lifeexp.txt", header = TRUE)
```

### a
```{r}
pairs(lifeexp, panel = panel.smooth)
```
Appears to be nonlinear, People per TV and Doctor both highly skewed.

### b
```{r}
lifeexp$logpptv = log(lifeexp$People.per.TV)
lifeexp$logppd = log(lifeexp$People.per.Dr)

head(lifeexp)
```
```{r}
pairs(LifeExp ~ logpptv + logppd, data = lifeexp,
panel = panel.smooth)
```
Structure in the data between the response and log predictors looks much more linear. We should be
able to fit a linear model now to this data.

### c
```{r}
life.1 <- lm(LifeExp ~ logpptv + logppd, data = lifeexp)

par(mfrow = c(2, 2))
plot(life.1, which = 1:2)
plot(resid(life.1) ~ logpptv + logppd, data = lifeexp)
```
Quantile plot looks close to linear, some divergence away from linear. So errors near close to normally
distributed. Residual plots have no pattern, suggesting linear model adequate.


### d
```{r}
summary(life.1)
```
Might have to come back for answer.

