---
title: "week3sgta"
author: "Al"
date: "2023-08-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1
## Question 1

We will conduct a two sample, two tailed unpaired t-test since the population SD is unknown and we're looking for a change in the means. The null hypothesis will assume equal means. However we should first check assumptions. 
- Normal distribution 
- Large enough sample size
- Equal variances
There's also the assumption of independance however this cannot be verified.
```{r}
light <- read.table("SGTADATASETS/light.dat", header = TRUE)

#Check normal distribution
par(mfrow = c(1, 2))
{qqnorm(light$Brand1)
qqline(light$Brand1)}
{qqnorm(light$Brand2)
qqline(light$Brand2)} 
#Both plots follow normal distribution, safe assumption

#Check whether variance is equal
c(sd(light$Brand1), sd(light$Brand2))
boxplot(light) 
#The distributions are pretty similar, safe assumption 
```

Since we have determined the assumptions, we can move forward with our calculations.

```{r}
t.test(light$Brand1, light$Brand2, var.equal = TRUE)
```

Since p-value = 0.037 < 0.05, we have evidence to reject the null hypothesis and so there is reason to believe the lumen levels are different between both brands. The mean of brand 1 is 80.3, and brand 2 is 67.4. We are 95% confident that the brightness level of brand 1 light bulbs is, on average, [0.84, 24.98] lumens brighter than brand 2. 

## Question 2
We will conduct a paired t-test as the two plots are related. We will conduct a t-test on the differences with null hypothesis mu_d = 0, and alternative hypothesis my_d > 0. This will require a two tailed t-test. We should first check assumptions:
- Independence
- Normality of differences

```{r}
farm <- read.table("SGTADATASETS/farm.dat", header = TRUE)
difference = farm$yield[farm$plot == "sprayed"] - farm$yield[farm$plot == "unsprayed"]

#Check normality
boxplot(difference)
{qqnorm(difference) 
qqline(difference)}
```
The data shows there is a deviation in the lower tail area. We can exclude an outlier however the deviation remains. We will continue assuming normality despite this.

```{r}
t.test(difference, alternative = 'g')
#t.test(yield ~ plot, paired = TRUE, data = farm, alternative = 'g')
```
We have found that p-value of the differences = 0.022 < 0.05. We have evidence to reject the null hypothesis and believe that the true mean of the differences is greater than 0 i.e. spraying increases crop yields. We are 95% confident that, on average, spraying yields crops that are greater by 0.56 units.

## Question 3
Since the gravel levels are from two different types of soil selected at random, we can apply the two sample t-
test to compare the mean parameters in fine gravel levels. This test will be valid so long as the measurements
are independent of each other and the measurements are close to normally distributed. There is no preference
as to the difference so we do a two-sided test.

```{r}
fertile <- read.table("SGTADATASETS/fertile.dat", header = TRUE)

#Check normality
par(mfrow = c(1,2))
{qqnorm(fertile$fertile)
  qqline(fertile$fertile)}
{qqnorm(fertile$nonfertile)
  qqline(fertile$nonfertile)}

#Check variance
c(sd(fertile$fertile), sd(fertile$nonfertile))
boxplot(fertile)
```
While there is evidence of normality, the boxplots show too large of a difference in variances, so we must operate on the assumption
of unequal variances.
```{r}
t.test(fertile$fertile, fertile$nonfertile, ver.equal = FALSE)
```
Since our P-Value is 0.21294 > α = 0.05, we do not have evidence against H0 suggesting that the
means are not significantly different for the two soil types. That is, we don’t have evidence to suggest
the different soil types have different fine gravel levels.



