---
title: "week8sgta"
author: "Al"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1
## Question 1
### a
```{r}
ecoli = read.table("SGTADATASETS/ecoli.txt", header = TRUE)
str(ecoli)
{plot(HEC ~ HGMF, data = ecoli)
  abline(lm(HEC ~ HGMF, data = ecoli))}
```

### b)
```{r}
ecoli.lm = lm(HEC ~ HGMF, data = ecoli)
summary(ecoli.lm)
```
### c)
```{r}
plot(ecoli.lm, which = 1:2)
```
The residual plot has a large outlier, which is also influencing the qqplot evident by its left skew.

### d)
There are two ways to identify the outlier:

#### 'Plotting' Way
We have already used this in the former plots. Conveniently the furthest points are labelled. In this case observation 18 is the outlier.

#### Manual Way
If the points weren't labelled, we can see that the outlier is close to -1. So we can call:
```{r}
badpoint = which(ecoli.lm$residuals < -0.5)
badpoint
```

### e)
```{r}
cleanEcoli = ecoli[-badpoint, ]

better.ecoli.lm = lm(HEC ~ HGMF, data = cleanEcoli)
plot(better.ecoli.lm, which = 1:2)
```
The quantile plot and residual plot look much better once the outlier is removed, there is a nice linear
trend in the Q-Q plot and no obvious pattern or outlier in the residual plot. Plotting the new analysis
and adding the prediction and confidence intervals.

```{r}
plot(HEC ~ HGMF, data = cleanEcoli)
abline(better.ecoli.lm)
# Create sequence of points
hgmfgrid <- seq(from = -1, to = 3, by = 0.5)
# Put in a dataframe with the same name as the predictor (HGMF)
HGMFdat <- data.frame(HGMF = hgmfgrid)
# Create the intervals
predvals <- predict.lm(better.ecoli.lm, newdata = HGMFdat, interval = 'prediction')
confvals <- predict.lm(better.ecoli.lm, newdata = HGMFdat, interval = 'confidence')
# Add prediction bands to plot, lty = 2 creates dashed line
# Note lty is short for 'line type'
lines(hgmfgrid, predvals[, 2], lty = 2)
lines(hgmfgrid, predvals[, 3], lty = 2)
# Add confidence bands to plot, lty = 3 creates dotted line
lines(hgmfgrid, confvals[, 2], lty = 3)
lines(hgmfgrid, confvals[, 3], lty = 3)
```


```{r}
bandind = which(hgmfgrid == 1)
bandind

predband1 = predvals[bandind, 2:3]
confband1 = confvals[bandind, 2:3]

predband1
confband1
```
## Question 2
```{r}
summary(better.ecoli.lm)
```
HGMF = 0.91568, and the standard error is 0.04356. 
```{r}
qt(0.975, 15)
```
0.91568 ± 2.1314495 × 0.04356 = (0.82283, 1.00853)

## Question 3
IDK4now

## Question 4
```{r}
anova(better.ecoli.lm)

```

## Question 5
Generally speaking, the model assumes that the predictor is measured without error. On paper, that means
we should use the more accurate variable as the predictor. Here we are assuming that as HGMF is a more
complex method that it should have a lower error (otherwise what’s the point of having this new method!).
In this type of problem, other techniques are required to obtain a predicted HGMF from HEC, see section
11.6 of Ott and Longnecker, An Introduction to Statistical Methods and Data Analysis.
In reality, there are other factors to consider, such as the cost and ease to obtain the predictor and response.
Things are often less clear in most applied studies, and it is up to you to make a case for it
